---
title: "Thematic Mapping with tmap"
author: "Maria Kamenetsky"
date: "January 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

The `R` package `tmap` includes extensive functionality for making quality maps. We will use it here. 

# Set-Up

```{r}
library(tmap)
library(dplyr)
library(sf)
```


Import shape file of class `sf` (created in *Weighting and Transformations* tutorial):

```{r}
pv <- read_sf(dsn="../data", layer="pov_uscounties_sf")
class(pv)
```

For the example, we will only look at a single state, Wisconsin (state FIPS code 55):

```{r}
wisc <- pv %>%
    filter(STATEFP=="55")
str(wisc)
```




# Thematic Mapping

```{r}
#basic grey map with county borders
basemap <- tm_shape(wisc) +
    tm_borders() 
basemap + tm_fill() 
```


We can also plot one of the variables, *poverty*:

```{r}
povmap <- basemap + tm_fill(col="poverty")
povmap
```

Create a map with panels of maps:

```{r}
#create and store maps
agmap <- basemap + tm_fill(col="ag")
retailmap <- basemap + tm_fill(col="retail")
manumap <- basemap + tm_fill(col="manu")    

#those maps into panels
tmap_arrange(povmap, agmap, retailmap, manumap)
```


Make an interactive map you can scroll over and visually explore poverty in Wisconsin:

```{r}
tmap_mode("view")
basemap + tm_polygons(col="poverty")
```


```{r}
tmap_mode("plot")
basemap + tm_polygons(col="poverty") + 
    tm_symbols(col = "red", size = "totpopn", alpha=0.5) +
    tm_layout(legend.title.size = 0.7,
          legend.text.size = 0.5,
          legend.position = c("right","top"),
          legend.bg.color = "white",
          legend.bg.alpha = 1)
```



Add a point to Madison, Wisconsin:

```{r}
mad <- data.frame(c("Madison,Wisconsin"),
                  lat=43.0731,
                  long = -89.4012)
site_mad <- sf::st_as_sf(mad, coords = c("long", "lat"),
                         crs = "+proj=longlat +datum=WGS84",
                         agr="identity")

basemap + tm_shape(site_mad) + 
    tm_dots(size=2, col="red", shape=21) +
    tm_compass() +
    tm_layout(title="Madison, WI", title.position = c("center","top"))

```




